\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%\renewcommand{\baselinestretch}{2}
\pdfminorversion=4
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}



% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\input{macros}
\def\fat#1{#1}

\title{Transfer Entropy MDPs with  Temporal Logic Specifications
}

\author{Suda Bharadwaj \and Mohamadreza Ahmadi \and
 Takashi Tanaka \and Ufuk Topcu% <-this % stops a space
%\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{All authors are with the University of Texas at Austin. E-mail: \{suda.b, mrahmadi, ttanaka, utopcu\}@utexas.edu}%
%\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University, Dayton, OH 45435, USA {\tt\small b.d.researcher@ieee.org}}%
}


\tikzstyle{block} = [draw, fill=blue!20, rectangle, 
    minimum height=3em, minimum width=6em]
\tikzstyle{sblock} = [draw, fill=blue!20, rectangle, 
    minimum height=3em, minimum width=3em]
\tikzstyle{sum} = [draw, fill=blue!20, circle, node distance=1.2cm]
\tikzstyle{input} = [coordinate]
\tikzstyle{output} = [coordinate]
\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]



\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
Applications in autonomy, in particular, space rovers motivate the need for control techniques that take into account uncertain environments, communication, and sensing constraints, while satisfying high-level mission specifications. In order to address such  challenges, we consider a class of Markov decision processes (MDPs), where the cost function is \emph{transfer entropy} which we refer to as TEMDPs. In this context, we study high-level mission specifications in terms of co-safe linear temporal logic (LTL) formulae.   Based on the dynamic programming principle, we provide a method to synthesize a policy that minimizes the weighted sum of the cost function and the probability of failure to satisfy the LTL specification. We show that this policy can be obtained using a modified Arimoto-Blahut algorithm. Finally, we show the efficacy of the proposed method by applying it to address navigation and path planning scenarios of a Mars rover, demonstrating the effect of considering the transfer entropy~cost.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\input{intro}

%\section{Case Study}\label{sec:casestudy}
%\input{casestudy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Preliminaries}
The sequence $(x_0,x_{1}...x_t)$ is denoted $x^t$ and the subsequence $x_l,x_{k+1}...x_k$ is denoted by $x_{l}^{k}$. We use upper-case letters to denote random variables and lower-case letters for the realizations of the corresponding random variable. \\
We denote by $\dist{\mathcal{X}}$ the set of all probability distributions on a finite
set $\mathcal{X}$, \ie all functions $f: \mathcal{X} \to [0,1]$ such that $\sum_{x\in \mathcal{X}}f(x)=1$. Finally, for a set $\mathcal{S}$, we define $2^\mathcal{S}$ as the set of all subsets of $\mathcal{S}$ and $S^{\omega}$ as the set of all infinite sequences of elements in $\mathcal{S}$
\subsection{Markov decision processes}
\input{MDP} 

\subsection{Temporal Logic}
\input{LTL}

\subsection{Information cost}
\input{infotheory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Problem statement}
\input{problem_Takashi_2}

\section{Incorporating temporal logic constraints}
\input{problem_Suda}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Optimality conditions}
\input{opt_cond_tak}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\subsection{Arimoto-Blahut algorithm}
\input{AB_alg}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Interpretation of directed information}\label{sec:ab}
%\input{comm_rate.tex}


\section{Experiment results}\label{sec:exp}
\input{case_study}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion and future work}
In this paper, we presented a formal way to integrate co-safe LTL constraints into a minimal-information MDP problem. This is the first step in analyzing temporal logic constraints in communication constrained problems. For future work, we aim to relax the co-safe requirement to allow more general classes of LTL formulas by analyzing the mean information cost over an infinite run. Furthermore, we aim to extend this work to a multiple coordinating agent formulation as this problem setting naturally lends itself to minimizing communication between agents who are trying to satisfy a joint specification. 



\bibliographystyle{IEEEtran}
\bibliography{main}

\appendices
\vspace{-0.15cm}
%\section{Proof of Theorem 5.1} \label{sec:prf}
%\input{app1_off}

\end{document}
